### 2012-11-02 icse2013demo-testevol/paper.tex
Minor fix

\documentclass[conference]{IEEEtran}
\synctex=1

\usepackage{times}
\usepackage{graphicx}
\usepackage{epsfig}
% \usepackage{url}
% \usepackage{hyperref}
\usepackage[bookmarks=false]{hyperref}
\usepackage{xspace}
\usepackage{float}
\usepackage{latexsym}
\usepackage{natbib}
\usepackage[usenames]{color}
\usepackage{multirow}
\usepackage{listings}
\usepackage{textcomp}

\bibpunct{[}{]}{,}{n}{}{}

\def\denseitems{
  \itemsep1pt plus1pt minus1pt
  \parsep0pt plus0pt
  \parskip0pt\topsep0pt}

\input{mymacros}

\makeatletter
\newcommand\tabcaption{\def\@captype{table}\caption}
\makeatother

\renewcommand\lstlistingname{\small{Listing}}

\newcommand{\lang}[1]{\texttt{\small #1}}
\newcommand{\subject}[1]{\texttt{\small #1}}
\newcommand{\mt}{\mathit}
\newcommand{\trex}{\textsc{TestEvol}}
\newcommand{\pass}{\mt{Pass}}
\newcommand{\fail}{\mt{Fail}}
\newcommand{\failce}{\mt{Fail}_{CE}}
\newcommand{\failre}{\mt{Fail}_{RE}}
\newcommand{\failae}{\mt{Fail}_{AE}}
\newcommand{\testfunc}[2]{\mt{Test(#1, #2)}}
\newcommand{\covfunc}[2]{\mt{Cov(#1, #2)}}
% \newcommand{\note}[1]{{\color{red}$\blacktriangleright$ \bf #1
%     $\blacktriangleleft$}}

\newcommand{\catrep}{\textsc{TestRep}}
\newcommand{\catref}{\textsc{TestModNotRep}}
\newcommand{\catdelaere}{\textsc{TestDel}$_\mt{(AE|RE)}$}
\newcommand{\catdelce}{\textsc{TestDel}$_\mt{(CE)}$}
\newcommand{\catdelp}{\textsc{TestDel}$_\mt{(P)}$}
\newcommand{\cataddaere}{\textsc{TestAdd}$_\mt{(AE|RE)}$}
\newcommand{\cataddce}{\textsc{TestAdd}$_\mt{(CE)}$}
\newcommand{\cataddp}{\textsc{TestAdd}$_\mt{(P)}$}

\newcommand{\smcatrep}{{\small \textsc{TestRep}}}
\newcommand{\smcatref}{{\small \textsc{TestRef}}}
\newcommand{\smcatdelaere}{{\small \textsc{TestDel}$_\mt{(AE|RE)}$}}
\newcommand{\smcatdelce}{{\small \textsc{TestDel}$_\mt{(CE)}$}}
\newcommand{\smcatdelp}{{\small \textsc{TestDel}$_\mt{(P)}$}}
\newcommand{\smcataddaere}{{\small \textsc{TestAdd}$_\mt{(AE|RE)}$}}
\newcommand{\smcataddce}{{\small \textsc{TestAdd}$_\mt{(CE)}$}}
\newcommand{\smcataddp}{{\small \textsc{TestAdd}$_\mt{(P)}$}}

\lstset{ %
upquote=true,
basicstyle=\tiny\sffamily,
showspaces=false,               % show spaces adding particular underscores
  showstringspaces=false,         % underline spaces within strings
  showtabs=false
}

%\linespread{0.97}
%\columnsep 0.11in
\clubpenalty = 10000
\widowpenalty = 10000
\displaywidowpenalty = 10000

\hyphenation{Test-Evol}

\newcommand{\tool}{\textsc{TestEvol}\xspace}

\let\oldthebibliography=\thebibliography
\let\endoldthebibliography=\endthebibliography
\renewenvironment{thebibliography}[1]{%
  \begin{oldthebibliography}{#1}%
    \setlength{\parskip}{0ex}%
    \setlength{\itemsep}{0ex}%
  }%
  {%
  \end{oldthebibliography}%
}

%\toappear{}

\begin{document}

\title{TestEvol: A Tool for Analyzing Test-Suite Evolution}

\author{\IEEEauthorblockN{XXX}
\IEEEauthorblockA{
% College of Computing\\
Georgia Institute of Technology\\
% Atlanta, USA\\
\{orso\}@cc.gatech.edu}
}

\maketitle

\pagestyle{empty}

\note{This is a VERY rough draft that I put together by copying and
  pasting from the FSE paper. I left some specific TODOs in, but the
  paper really needs a major pass.}

\todo{Restructure the content to make sure we comply with what is
  requested from a demo paper: Tool-based demonstrations describe
  novel aspects of early prototypes or mature tools. The tool
  demonstrations must communicate clearly the following information to
  the audience:\\
  - the envisioned users;\\
  - the software engineering challenge it proposes to address;\\
  - the methodology it implies for its users; and\\
  - the results of validation studies already conducted for mature tools, or the design of planned studies for early prototypes.}

\todo{Reduce the description of the approach and extend the
  description of the implementation. Leandro, can you add some details
  there?}

\todo{Add snapshots.}

\begin{abstract}
  Test suites, once created, rarely remain static. Just like the
  application they are testing, they evolve throughout their lifetime.
  Test obsolescence is probably the most known reason for test-suite
  evolution---test cases cease to work because of changes in the code
  and must be suitably repaired. Repairing existing test cases
  manually, however, can be extremely time consuming, especially for
  large test suites, which has motivated the recent development of
  automated test-repair techniques. We believe that, for developing
  effective repair techniques that are applicable in real-world
  scenarios, a fundamental prerequisite is a thorough understanding of
  how test cases evolve in practice. Without such knowledge, we risk
  to develop techniques that may work well for only a small number of
  tests or, worse, that may not work at all in most realistic cases.
  Unfortunately, to date there are no studies in the literature that
  investigate how test suites evolve. To tackle this problem, in this
  demo we present \tool, a tool that enables the systematic study of
  test-suite evolution for Java programs and JUnit test cases. We also
  present examples of applications of \tool\ to real-world programs
  and their unit test suites that show the usefulness and practical
  applicability of the tool.
\end{abstract}

% A category with the (minimum) three required fields
%\category{H.4}{Information Systems Applications}{Miscellaneous}
%A category including the fourth, optional field follows...
%\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]

%\terms{Theory}

%\keywords{ACM proceedings, \LaTeX, text tagging}

\section{Introduction}
\label{sec:intro}

Test cases form the first line of defense against the introduction of
software faults, especially when retesting modified software.  With
the availability of convenient testing frameworks, such as JUnit, and
the adoption of agile development methodologies, in particular,
writing unit test cases is an increasingly common practice nowadays.
Developers routinely create test suites that consist of unit tests and
that they run periodically on their code.

Test suites are not static entities: they constantly evolve along with
the application they test.  In particular, changes in the application
can break test cases---in some cases, even a small change in the
application code can affect a large number of tests. In general, a
test failure on a new version of the software can either expose
application faults or result from a problem with the test itself.  On
observing a failure, the first task for the developer is to determine
which of these two is the cause of the failure and, once this
distinction is made, either fix the problem in the code or fix the
broken~test.

A broken test, if it covers a valid functionality, should ideally be
repaired. Alternatively, if the repair is unduly complex to perform or
if the test was designed to cover a functionality that no longer
exists in the application, the test should be removed from the test
suite. To illustrate with an example,
Figure~\ref{fig:pmd-javatokenizertest} shows two versions of a unit
test case from the test suite of \subject{PMD}, one of the programs
used in our demo. A change in \subject{PMD}'s API broke the original
version of the test case, which had to be fixed by adding a call to
method \lang{SourceCode.readSource} and removing one parameter from
the call to method \lang{Tokenizer.tokenize} (lines 6 and 7 in
Figure~\ref{fig:pmd-javatokenizertest}(b), respectively).

\begin{figure}[t]
\lstinputlisting[language=java,
captionpos=t,frame=tb,stepnumber=1,numbers=left,numbersep=5pt]{code/a.java}
\vspace*{-4pt}
\centerline{(a)}
\lstinputlisting[language=java,
captionpos=t,frame=tb,stepnumber=1,numbers=left,numbersep=5pt]{code/b.java}
\vspace*{-4pt}
\centerline{(b)}
\vspace*{-8pt}
\caption{Two versions of a test case from \subject{PMD}'s unit test
  suite: (a) version~1.4, broken, and (b) version~1.6, repaired.}
\vspace*{-12pt}
\label{fig:pmd-javatokenizertest}
\end{figure}

Because test repair can be an expensive activity, automating it---even
if only partially---could save a considerable amount of resources
during maintenance. This is the motivation behind the development of
automated test-repair techniques, such as the ones targeted at unit
test cases~\cite{Daniel:2009, Daniel:2010, Mirzaaghaei:2012} and those
focused on GUI (or system) test cases~\cite{Choudhary:2011,
  Grechanik:2009, Huang:2010, Memon:2008}.

We believe that, to develop effective techniques for assisting manual
test repair, we must first understand how test suites evolve in
practice. That is, we must understand when and how tests are created,
removed, and modified. This is a necessary, preliminary step because
it can (1) provide evidence that test cases do get repaired, (2)
support the hypothesis that test repairs can be (at least partially)
automated, and (3) suitably direct research efforts.  Without such
understanding, we risk to develop techniques that may not be generally
applicable and may not perform the kind of repairs that are actually
needed in real-world software systems.

To the best of our knowledge, to date there are no studies in the
literature that investigate how unit test suites evolve. To address
this issue, we defined a technique~\cite{pinto12} that combines
various static- and dynamic-analysis techniques to compute the
differences between the test suites associated with two versions of a
program and categorize such changes along two dimensions: the static
differences between the tests in the two test suites and the
behavioral differences between such tests.  We implemented our
technique in a tool, \tool, which enables a systematic study of
test-suite evolution. \tool\ targets Java programs and JUnit test
suites, as Java is a widely used language and JUnit is the de-facto
standard unit-testing framework for Java.  Given two versions of a
program and its test suite, \tool\ automatically computes differences
in the behavior of the test suites on the two program versions,
classifies the actual repairs performed between the versions, and
computes the coverage attained by the tests on the two program
versions.

In this demo, we present \tool, its features and technical details,
and show how it can be used to study how test suites evolved over the
years for a project.  To do so, we will show examples of application
of the tool to a set of real-world open-source software systems.
Specifically, we will show how to use \tool\ to investigate relevant
questions on test evolution, such as: What types of test-suite changes
occur in practice and with what frequency? How often do test repairs
require complex modifications of the tests? Why are tests deleted and
added? Overall, we will show how \tool\ is a useful tool not only for
researchers in the area of test repair (and test evolution in
general), but also to developers and testers who want to better
understand their software systems.

\todo{Discuss \tool's availability at
  \url{http://cheetah.cc.gt.atl.ga.us:8081/testevol/}}.

\section{Our Approach for Studying Test-Suite Evolution}
\label{sec:test-evolution}
\subsection{Definitions and Terminology}
\label{sec:terminology}

A \textit{system} $S = (P, T)$ consists of a program $P$ and a test
suite $T$.  A \textit{test suite} $T = \{t_1, t_2, \ldots, t_n\}$
consists of a set of unit test cases. 
% A unit test case $t$ is a method (or function) and it is identified
% by its name plus the name of the class it is associated with.
$\testfunc{P}{t}$ is a function that executes test case $t$ on program
$P$ and returns the outcome of the test execution.  A \textit{test
  outcome} can be of one of four types:

\begin{itemize}\denseitems

\item $\pass$: The execution of $P$ against $t$ succeeds.

\item $\failce$: The execution of $P$ against $t$ fails because a
  class or method accessed in $t$ does not exist in
  $P$.\footnote{\small These failures can obviously be detected at
    compile-time. For the sake of consistency in the discussion,
    however, we consider such cases to be detected at runtime via
    ``class not found'' or ``no such method'' exceptions. In fact, our
    \tool{} tool detects such failures at runtime by executing the
    tests compiled using the previous version of $P$ on $P$.}

\item $\failre$: The execution of $P$ against $t$ fails due to an
  uncaught runtime exception (\eg a ``null pointer'' exception).

\item $\failae$: The execution of $P$ against $t$ fails due to an
  assertion violation.

\end{itemize}

We use the generic term $\fail$ to refer to failures for which the
distinction among the above three types of failures is unnecessary.

$\covfunc{P}{t}$ is a function that instruments program $P$, executes
test case $t$ on $P$, and returns the set of all statements in $P$
covered by $t$. $\covfunc{P}{T}$ returns the cumulative coverage
achieved on $P$ by all the tests in test suite $T$.

% $\covfunc{P}{t}$ is a function that instruments program $P$, executes
% test case $t$ on $P$, and returns the number of branches in $P$
% covered by $t$. $\covfunc{P}{T}$ returns the cumulative coverage
% attained on $P$ by all the tests in test suite $T$.

%% Unit tests are run very frequently---often from the IDE, soon after
%% the code is changed, or even while the code is being changed. In a
%% typical testing process, developers may run the test suite after
%% editing the code. In the run, some tests may fail, whereas others
%% would pass. The developer would want to fix the broken tests online to
%% ensure that all tests pass.

Given a system $S = (P, T)$, a modified version of $S$, $S'=(P', T')$,
and a test case $t$ in $T \cup T'$, there are three possible scenarios
to consider: (1) $t$ exists in $T$ and $T'$, (2) $t$ exists in $T$ but
not $T'$ (\ie $t$ was removed from the test suite), and (3) $t$ exists
in $T'$ but not in $T$ (\ie $t$ was added to the test suite). These
scenarios can be further classified based on the behavior of $t$ in
$S$ and $S'$, as summarized in Figure~\ref{fig:study-design} and
discussed in the rest of this section.

\subsection{Test Modifications}
\label{sec:test-mod}

Figure~\ref{fig:study-design}(a) illustrates the scenario in which $t$
is present in the test suites for both the old and the new versions of
the system. To study different cases, we consider whether $t$ is
modified (to $t'$) and, if so, whether the behaviors of $t$ and $t'$
differ. (We have not considered the cases in which $t$ is not modified
because they are irrelevant for studying test evolution.)  For
\textit{behavioral differences}, there are two cases, shown in the two
rows of the table: either $t$ fails on $P'$ and $t'$ passes on $P'$ or
both $t$ and $t'$ pass on $P'$.

\begin{figure}[t]
\centering
\small
\tabcolsep=3pt
\textbf{(a) Test $t$ exists in $S$ and $S'$ and is modified}
\\ [2pt]
\begin{tabular}{|l||l|}
\hline
%\multicolumn{1}{|c||}{} & \multicolumn{1}{|c}{} &
%\multicolumn{1}{|c|}{$t$ is not} \\
%\multicolumn{1}{|c||}{} & \multicolumn{1}{|c|}{$t$ is modified}\\
%\hline \hline
$\testfunc{P'}{t} = \fail \; \wedge$ &
$t$ is repaired \\
$\testfunc{P'}{t'} = \pass$ &
\multicolumn{1}{r|}{\scriptsize [\catrep{}]}\\
\hline
\multirow{3}{28mm}{$\testfunc{P'}{t} = \pass \; \wedge \testfunc{P'}{t'} = \pass$} &
$t$ is refactored, updated to test a different \\
& scenario, or is made more/less discriminating \\
& \multicolumn{1}{r|}{\scriptsize [\catref{}]} \\
\hline
\end{tabular}
\\ [8pt]
%
\textbf{(b) Test $t$ is removed in $S'$}
\\ [2pt]
\begin{tabular}{|l||l|}
\hline
\multirow{2}{*}{$\testfunc{P'}{t} = \failre \, | \, \failae$} &
$t$ is too difficult to fix \\
& \multicolumn{1}{r|}{\scriptsize [\catdelaere{}]} \\
\hline
\multirow{2}{*}{$\testfunc{P'}{t} = \failce$} & $t$ is obsolete or is
too difficult to fix \\
& \multicolumn{1}{r|}{\scriptsize [\catdelce{}]} \\
\hline
\multirow{2}{*}{$\testfunc{P'}{t} = \pass$} & $t$ is redundant \\
& \multicolumn{1}{r|}{\scriptsize [\catdelp{}]} \\
\hline 
\end{tabular}
\\ [8pt]
%
\textbf{(c) Test $t'$ is added in $S'$}
\\ [2pt]
\begin{tabular}{|l||l|}
  \hline
  \multirow{2}{*}{$\testfunc{P}{t'} = \failre \, | \, \failae$} &
  $t'$ is added to validate a bug fix \\
  & \multicolumn{1}{r|}{\scriptsize [\cataddaere{}]} \\
  \hline
  \multirow{3}{*}{$\testfunc{P}{t'} = \failce$} &
  $t'$ is added to test a new functionality \\
  & or a code refactoring \\
  & \multicolumn{1}{r|}{\scriptsize [\cataddce{}]} \\
  \hline
  \multirow{3}{*}{$\testfunc{P}{t'} = \pass$} &
  %% TODO: Check: $t'$ is added to test a new functionality \\
  $t'$ is added to test an existing feature\\
  & or for coverage-based augmentation \\
  & \multicolumn{1}{r|}{\scriptsize [\cataddp{}]} \\
  \hline
\end{tabular}
\caption{Scenarios considered in our investigation. Given two system
  versions $S = (P, T)$ and $S' = (P', T')$, the three scenarios are:
  (a) $t$ exists in $T$ and $T'$ and is modified, (b) $t$ exists in
  $T$ but not in $T'$, (c) $t'$ exists in $T'$ but not in $T$.}
\vspace*{-5pt}
\label{fig:study-design}
\end{figure}

\subsubsection{Category \catrep{}: Repaired Tests}
\label{sec:category-rep}

The \catrep{} category corresponds to the case where $t$ is repaired
so that, after the modifications, it passes on $P'.$ As discussed in
the Introduction, Figure~\ref{fig:pmd-javatokenizertest} shows an
example of such a test repair. The code fragments in
Listings~\ref{c1-old} and~\ref{c1-new} present another example of
repair that involves a simpler code modification than the one in
Figure~\ref{fig:pmd-javatokenizertest}. The example is taken from
\subject{Gson}, one of the programs used in our demo, and involves a
test case from \subject{Gson} version~2.0 that was fixed in the
subsequent version~2.1. Test \lang{testNullField} had to be fixed
because constructor \lang{FieldAttributes(Class<?> declClazz, Field
  f)} from version~2.0 was modified in version~2.1 to take only one
parameter (of type \lang{Field}).
%% was no longer available on version 2.1. In the new version of the
%% \lang{FieldAttributes} class, the constructor receives only one
%% parameter of type \lang{Field}.

\vspace*{-10pt}
\lstinputlisting[language=java,label=c1-old,caption=\small{Unit test
  for class FieldAttributes} (Gson v2.0),
captionpos=t,frame=tb]{code/OldTestRep.java}

\vspace*{-5pt}
\lstinputlisting[language=java,label=c1-new,caption=\small{Unit test
  for class FieldAttributes} (Gson v2.1),
captionpos=t,belowcaptionskip=4pt,frame=tb]{code/TestRep.java}

For this category, we wish to study the types of modifications that
are made to $t$. A test repair may involve changing the sequence of
method calls, assertions, data values, or control flow. Based on our
experience, for \textit{method-call sequence changes}, we consider
five types of modifications:

\begin{enumerate}
\denseitems

\item \textit{Method call added}: a new method call is added.

\item \textit{Method call deleted}: an existing method call is
  removed.

\item \textit{Method parameter added}: a method call is modified such
  that new parameters are added.

\item \textit{Method parameter deleted}: a method call is modified
  such that existing parameters are deleted.

\item \textit{Method parameter modified}: a method call is modified
  via changes in the values of its actual parameters.

\end{enumerate}

A test repair may involve multiple such changes. For example, the
repair shown in Figure~\ref{fig:pmd-javatokenizertest} involves the
addition of a method call~(in~line~6) and the deletion of a method
parameter (in line~7). The repair illustrated in Listings~\ref{c1-old}
and~\ref{c1-new} also involves the deletion of a method parameter.
%
For \textit{assertion changes}, we consider cases in which an
assertion is added, an assertion is deleted, the expected value of an
assertion is modified, or the assertion is modified but the expected
value is unchanged.
%
Finally, we currently group together \textit{data-value changes} and
\textit{control-flow changes}.

The rationale underlying our classification scheme is that different
classes of changes may require different types of repair analyses.
Although this is not necessarily the case, at least the search
strategy for candidate repairs would differ for the different
categories of changes. Consider the case of method-parameter deletion,
for instance, for which one could attempt a repair by simply deleting
some of the actual parameters. Whether this repair would work depends
on the situation. For the code in
Figure~\ref{fig:pmd-javatokenizertest}, for example, it would not work
because the deletion of one of the parameters in the call to
\lang{tokenize()} is insufficient by itself to fix the test---a new
method call (to \lang{readSource()}) has to be added as well for the
test to work correctly.  Similarly, for the case of method-parameter
addition, one could conceivably attempt straightforward fixes by
constructing equivalence classes for the new parameters and selecting
a value from each equivalence class (\eg positive, negative, and zero
values for an integer parameter). In our study, we found cases where
such an approach would in fact repair broken tests. In this case too,
however, such a solution is in general not enough.

\subsubsection{Category \catref{}: Refactored Tests}
\label{sec:category-c2}

The \catref{} category captures scenarios in which a test $t$ is
modified in $S'$ even though $t$ passes on $P'$.
Listings~\ref{c2-old} and~\ref{c2-new} show an example of one such
change from \subject{Commons Math}. Unit test \lang{testDistance} was
refactored to invoke method \lang{sqrt()} in class \lang{FastMath}, a
newly added class in the new release, instead of the same method in
\lang{java.lang.Math}.  There are different reasons why changes in
this category might occur.

\vspace{-8pt}
\lstinputlisting[language=java,label=c2-old,caption=\small{Unit test
  from class Vector3DTest} (Commons Math v2.1),
captionpos=t,frame=tb]{code/OldTestRef.java}

\vspace{-8pt}
\lstinputlisting[language=java,label=c2-new,caption=\small{Unit test
  from class Vector3DTest} (Commons Math v2.2),
captionpos=t,frame=tb]{code/TestRef.java}

\subsection{Test Deletions}
\label{sec:test-ref}

Figure~\ref{fig:study-design}(b) illustrates the scenario in which a
test $t$ is deleted. To study the reasons for this, we examine the
behavior of $t$ on the new program version $P'$ and consider three
types of behaviors.

\subsubsection{Category \catdelaere{}: Hard-To-Fix Tests}
\label{sec:category-delaere}

%%COMMENT PLEASE READ ME: I guess here we should focus more
%%in the cases that tests are removed because they have become
%%obsolete. Actually, in the examples I was not able to find one
%%case in which it was removed because it was too difficult to
%% be fixed. It was always obsolete tests.

This category includes tests that fail on $P'$ with a runtime
exception or an assertion violation. These may be instances where the
tests should have been fixed, as the functionality that they test in
$P$ still exists in $P'$, but the tests were discarded instead. One
plausible hypothesis is that tests in this category involve repairs of
undue complexity, for which the investigation of new repair techniques
to aid the developer might be particularly useful.  We performed a
preliminary investigation of this hypothesis by manually examining ten
randomly selected tests in this category and found that all the
examined tests were in fact obsolete.

\subsubsection{Category \catdelce: Obsolete Tests}
\label{sec:category-delce}

A test that fails with a compilation error on the new program version
is obsolete because of API changes. Listing~\ref{c4-example}
illustrates a test in this category taken from \subject{JodaTime}.
This test was deleted because the tested method
\lang{Chronology.getBuddhist()} was removed in the subsequent version
of \subject{JodaTime}.

% \vspace{-4pt}
\lstinputlisting[language=java,label=c4-example,caption=\small{Unit
  test from class TestChronology} (JodaTime v2.0),frame=tb]
{code/TestDel-Ce.java}
%This method was
%already marked as deprecated in the previous version.

Although for this category of deletion too, one could postulate that
the tests were removed because they were too difficult to fix, we
believe this not to be the case in most practical occurrences.
Instead, the more likely explanation is that the tests were removed
simply because the tested methods were no longer present. Also in this
case, we investigated our hypothesis by manually examining ten
randomly selected cases. Indeed, our manual investigation confirmed
that, for the cases we analyzed, the tested functionality was either
removed or provided through alternative methods, which requires the
development of new tests rather than fixes to the existing ones.

\subsubsection{Category \catdelp: Redundant Tests}
\label{sec:category-delp}

This category includes tests that are removed even though they pass on
$P'$.  Listing~\ref{c5-example} illustrates an example of one such
test, taken from \subject{Commons Lang}. In the new version, this test
was replaced by a more sophisticated one, shown in
Listing~\ref{c5.2-example}.

\begin{figure}[h]
\lstinputlisting[language=java,label=c5-example,caption=\small{Unit
  test from class MatrixIndexExceptionTest} (Commons Lang
v2.1),frame=tb] {code/TestDel-P.java}
\vspace{4pt}
\end{figure}

\lstinputlisting[language=java,label=c5.2-example,caption=\small{Unit
  test implemented to replace the removed one} (Commons Lang
v2.2),frame=tb] {code/TestDel-PII.java}

\subsection{Test Additions}
\label{sec:test-add}

Figure~\ref{fig:study-design}(c) illustrates the cases of test-suite
augmentation, where a new test $t'$ is added to the test suite. The
behavior of $t'$ on the old program can indicate the reason why it may
have been added.

\subsubsection{Category \cataddaere: Bug-Fix Tests}
\label{sec:category-addaere}

This category includes added tests that fail on $P$ with a runtime
exception or an assertion violation. In this case, the functionality
that the added test $t'$ was designed to test exists in $P$ but is not
working as expected (most likely because of a fault). The program
modifications between $P$ and $P'$ would ostensibly have been made to
fix the fault, which causes $t'$ to pass on $P'$. Thus, $t'$ is added
to the test suite to validate the bug fix. Listing~\ref{c6-example}
illustrates a test that fits this profile, as it was added to
\subject{Commons Math} version 2.1 to validate a bug fix ({\small
  \url{https://issues.apache.org/jira/browse/MATH-391}}). In the
previous version of \lang{ArrayRealVector}, the creation of a
zero-length vector results in a runtime exception.

% \begin{figure}[h]
\lstinputlisting[language=java,label=c6-example,caption=\small{\mbox{Unit
  test from class ArrayRealVectorTest (Commons Math v2.2)}},
captionpos=t,frame=tb]{code/TestAdd-AeRe.java}
% \end{figure}

\subsubsection{Category \cataddce: New-Features Tests}
\label{sec:category-addce}

The added tests in this category fail on $P$ with a compilation error,
which indicates that the API accessed by the tests does not exist in
$P$. Thus, the added test $t'$ is created to test new code in $P'$,
where the code could have been added as part of a refactoring or, more
likely, to add new functionality.  Listing~\ref{c7-example}
illustrates a test from \subject{JFreeChart} that covers a new
functionality: method \lang{getMinY()} did not exist in previous
versions of class \lang{TimeSeries}.

\lstinputlisting[language=java,label=c7-example,caption=\small{Unit
  test from class TimeSeriesTests} (JFreeChart v1.0.14),
captionpos=t,frame=tb]{code/TestAdd-Ce.java}

\subsubsection{Category \cataddp: Coverage-Augmentation Tests}
\label{sec:category-addp}

This category considers cases where the added test $t'$ passes on
$P$. Clearly, $t'$ would have been a valid test in the old system as
well. One would expect that the addition of $t'$ increases program
coverage (\ie $\covfunc{P'}{T'} \supset \covfunc{P'}{T' -
  \{t'\}}$). Moreover, if $t'$ covers different statements in $P$ and
$P'$ (assuming that there is a way of matching statements between $P$
and $P'$), the plausible explanation is that $t'$ was added to test
the changes made between $P$ and $P'$. However, if $t'$ covers the
same statements in both program versions, it would have been added
purely to increase code coverage (and not to test any added or
modified code).  Listing~\ref{c8-example} illustrates an added test
case that increases code coverage: the new test method
\lang{testFindRangeBounds} covers the case where the parameter of
method \lang{findRangeBounds()} is null.

\lstinputlisting[language=java,label=c8-example,caption=\small{Unit test
from class XYErrorRendererTests} (JFreeChart v1.0.14),frame=tb]
{code/TestAdd-P.java}

\section{Implementation of \tool}
\label{sec:implementation}

\todo{Add the discussion of clone analysis.}

\todo{In general, add more details.}

\todo{Discuss the availability of the tool.}

We implemented our approach for studying test suite evolution in a
tool, called \tool, which facilitates a systematic study of test-suite
evolution for Java programs and JUnit test suites.  \tool\ analyzes a
sequence of versions of a software system, where each version consists
of application code and test code. The versions can be actual releases
or internal builds.  \tool\ consists of five components, as
illustrated in the architecture diagram shown in
Figure~\ref{fig:trex}.

The \emph{compiler} component builds each system version and creates
two jar files, one containing the application classes and the other
containing the test classes.
%
The \emph{test-execution engine} analyzes each pair of system versions
$(S, S')$, where $S = (P, T)$ and $S' = (P', T')$.  First, it executes
$T$ on program $P$ and $T'$ on program $P'$ (\ie it runs the tests on
the respective program versions).  Then, it executes $T'$ on $P$ and
$T$ on $P'$. For each execution, it records the test outcome: $\pass,
\failce, \failae, or \failre$.
%
The \emph{differencing component} compares $T$ and $T'$ to identify
modified, deleted, and added tests. This component is implemented
using the \textsc{wala} analysis infrastructure for Java ({\small
  \url{wala.sourceforge.net}}).

The test outcomes, collected by the test-execution engine, and the
test-suite changes, computed by the differencing component, are then
passed to the \emph{test classifier}, which analyzes the information
about test outcomes and test updates to classify each update into one
of the eight test-evolution categories presented in
Section~\ref{sec:test-evolution}.  For each pair of broken test case
and its repaired version, the test classifier also compares the test
cases to identify the types of repair changes---different types of
method-sequence changes and assertion changes---as discussed in
Section~\ref{sec:category-rep}; this analysis is also implemented
using \textsc{wala}.

\tool\ performs a further step for the test cases in categories
\catdelp{}and \cataddp{}. For these tests, the test classifier
leverages the \emph{coverage analyzer} to compute the branch coverage
achieved by each test; this facilitates the investigation of whether
the deleted or added tests cause any variations in the coverage
attained by the new test suite.

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{architecture.png}
        \vspace*{-16pt}
	\caption{High-level architecture of \tool.}
        \vspace*{-8pt}
	\label{fig:trex}
\end{figure}

\vspace*{-8pt}
\section{Conclusion}
\label{sec:summary}

Test suites evolve throughout their lifetime and change together with
the applications they are testing. Modifying existing test cases
manually to adapt them to an evolving system, however, can be
extremely time consuming, especially for large test suites, which has
motivated the recent development of automated test-repair techniques.
To support the development of more effective repair techniques, we
developed \tool, a tool for systematic investigation of test-suite
evolution that can provide its users with a thorough understanding of
how test cases evolve in practice. In this demo we present \tool, its
main features, and its technical characteristics. We also demonstrate
how \tool\ can be run on real-world software systems and produce a
wealth of useful information, such as the types of test-suite changes
that occur in practice and their frequency and why tests are deleted
and added.  Overall, we show how \tool\ is a useful tool not only for
researchers in the area of test repair, and test evolution in general,
but also to developers and testers who want to better understand their
software systems.

% \section*{Acknowledgments}

% This work was supported in part by NSF awards CCF-0916605 and
% CCF-0964647 to Georgia Tech, and by funding from IBM Research and
% Microsoft Research.

\bibliographystyle{abbrv}
\bibliography{paper}

\end{document}